Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2024-04-07 manuscript/manuscript.tex
commented on simulation, application and conclusion

Using the \textsl{qqplotr} and \textsl{ggplot2} packages 
\citep{qqplotr, ggplot2},
we constructed Q-Q plots of the distribution of the p-values to see if they are
uniformly distributed. If the distribution of the p-values is uniform, this 
indicates that under the null hypothesis, our method will only reject the $H_0$
at a rate of $\alpha$.
A zoomed in plot for probabilities between 0 and 0.1 is
also provided, as
that is the most common range for significance levels $\alpha$. 
%We can also
%compute the rate of rejection $\#P < \alpha$ for different values of
%$\alpha$, as well as a confidence interval
%for this proportion.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width = \textwidth]{figures/zoom_normal}
  \caption{A Q-Q plot of the p-values displayed in Figure~\ref{fig:normal} zoomed in to 
  probabilities between 0 and
  0.1}
  \label{fig:zoom_normal}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/zoom_gamma}
  \caption{A Q-Q plot of the p-values displayed in Figure~\ref{fig:gamma} zoomed in to 
  probabilities between 0 and
  0.1}
  \label{fig:zoom_gamma}
\end{figure}

From Figures~\ref{fig:normal} and \ref{fig:gamma}, we can observe that 
although the extent to which the p-values appear to be 
completely aligned with the line
in the Q-Q plots is dependent on the sample size and the level of serial 
dependence.
A small sample size like $n = 100$ or $200$ seems adequate when Kendall's
$\tau$ is lower than 0.25. For Kendall's $\tau \geq 0.5$, a sample larger than
$n = 200$ seems necessary for the p-values to be normally distributed. For
Kendall's $\tau \geq 0.75$, a sample larger than $n = 800$ appears to be 
necessary, as the p-values are not aligned with the line in the Q-Q plots, even
in Figures~\ref{fig:zoom_normal} and \ref{fig:zoom_gamma}. This
is not necessarily a cause for concern, as (for normal margins)
a Kendall's $\tau$ corresponds to
a $\phi$ of 0.924, which is very high. Additionally, performance generally seems
better for negative $\tau$ values versus positive $\tau$ values of the same
strength. Results are very similar for normal and gamma margins.

Under $H_0$, the p-values for the non-parametric block bootstrap KS test are
uniformly distributed. This is an indication that our method 
holds its size and that we can expect the probability
of a type 1 error to be the significance level $\alpha$.

\subsection{Power}
We must also evaluate if this method works when $H_0$ is false. We can use
the block bootstrap KS test to test if some $X_t \sim F$ is generated from 
some other 
distribution $G$ with some unknown $\theta$. In this scenario, if the test is 
indeed powerful,
we would ideally want $\beta$ 
(or the probability of failed rejection under $H_A$) 
to be 0, but we
expect it to be generally low. Again, we must try the method with different
marginal distributions to ensure that it is robust.
We would then expect the distribution of p-values to be non-uniform, and the 
rate
of $\#P < \alpha$ to be high, meaning a higher density of low p-values.

Like we did to see if the test holds its size, we generated time series with 
marginal distributions $N(8, 8)$ and 
$\Gamma(8, 1)$, Kendall's $\tau \in \{-.75, -.5, -.25, 0, .25, .5, 
.75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. However, 
to evaluate the test's power, when $X_t \sim N(8, 8)$, we tested
that the marginal distribution family is from the Gamma family,
or
$X_t \sim \Gamma(\cdot \mid \alpha, \beta), \quad t = 1, \ldots, n$.
for some unspecified $\alpha$ and $\beta$. Because the support of $N(8, 8)$ is
$(-\infty, \infty)$, but the support of $\Gamma(8, 1)$ is $(0, \infty)$, we used
the \textsl{truncdist} package \citep{truncdist} to truncate the series at 
values less than or equal to 0.
When $X_t \sim \Gamma(8, 1)$, we tested that the 
marginal distribution is from the
Normal family, or 
$X_t \sim N(\cdot \mid \mu, \sigma^2), \quad t = 1, \ldots, n$
for some unspecified $\mu$ and $\sigma$. For the block bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method $1000$ times 
to obtain a p-value $p_r$ for each $r \in \{1, \ldots, 1000\}$.

Given some significance level $\alpha$, we can compute a rejection rate 
$q = \#\{p_r < \alpha\} / 1000$ for $r \in \{1, \ldots, 1000\}$.
Additionally, we can construct a 95\% confidence interval for 
this rate. Below are tables containing rejection rates showcasing the 
differences when simulation settings are changed.

@@

From 
Figure~\ref{fig:rr}, 
the rejection rates appear to get lower as $\tau$ increases in a
positive direction. Using a sample size $n \geq 800$ appears to be sufficient
when $\tau$ is not too high, but for $\tau \geq 0.75$, all sample sizes 
observed appeared to have less than desirable rejection rates. Results
do not appear to be to different when comparing gamma and normal 
margins, indicating that the method is robust to the marginal distribution.
The rejection rates for $n = 800$ are very 
high (almost 1), indicating that under $H_A$, the test is very powerful 
(low probability of failed rejection $\beta$) when a large enough
sample size is used.

\section{Real Data Analysis}
\label{sec:real}

Using the \textsl{tseries} package 
\citep{tseries}, 
we gathered closing stock prices for the S\&P 500 index. We
observed how the method performs using the last 
4 years of data: January 1st, 2020 to December 31st, 2023.
We approximated stock returns by taking the difference in the logarithm 
of the closing prices.
Using our method, we tested that the stock returns are normally
distributed and Student's $t$ distributed 
with degrees of freedom $v \in \{30, 20, 10, 5, 4, 3, 2, 1\}$.
We tested the same hypotheses using \citet{babu2004goodness}'s 
non-parametric basic bootstrap bias correction and 
\citet{zeimbekakis2022misuses}'s parametric bootstrap method.
We used $B = 10,000$ bootstrap 
replicates in this analysis.
The p-values for these tests are highlighted in 
Table~\ref{table:SP5004}.

Once again, utilizing the \textsl{tseries} package \citep{tseries}, we collected
closing stock prices for the S\&P 100 index. Employing the same methodology as
our study on the S\&P 500 index, we conducted our analysis. The resulting 
p-values for these tests are presented in Table~\ref{table:SP1004}.

@@

While differences in p-values among different methods are expected, our analysis
reveals numerous instances where conclusions at the 0.05 significance level
using our method differ from those obtained using the methods proposed by
\citet{babu2004goodness} and \citet{zeimbekakis2022misuses}.

@@

For the S\&P 100 index, our method demonstrates more instances where it reaches
different conclusions at the 0.05 significance level compared to other methods.
Our method fails to reject the hypothesis that the series conforms to a
Student's $t$ distribution with degrees of freedom $v = 5$, yielding a p-value
of 0.0873. In contrast, \citet{babu2004goodness}'s method yields a p-value of 
0.0368, and \citet{zeimbekakis2022misuses}'s method yields a p-value of 0.03. 
We have here
another instance where employing different methods leads to divergent
conclusions regarding the marginal distribution of the population under scrutin.

@@

We have also 
demonstrated possible applications of this method to financial data.
Through comparison with \citep{babu2004goodness}'s non-parametric
bootstrap bias correction and \citet{zeimbekakis2022misuses}'s parametric
bootstrap method, we have underscored the importance of accounting for serial
dependence in choosing an appropriate methodology.
This method holds relevance for studies aiming to assess if a serially dependent
time series adheres to a hypothesized distribution in scenarios where parameters
are unknown or unspecified. Future studies could investigate
the method's performance on different marginal distributions and
different dependence structures. Additionally, one could propose a 
similar method
for comparing two serially dependent methods. One could also apply the method
showcased
in this paper to fields like earthquake prediction and astronomy.

>>>>

Using the \textsl{qqplotr} and \textsl{ggplot2} packages 
\citep{qqplotr, ggplot2},
we constructed Q-Q plots of the distribution of the p-values to see if they are
uniformly distributed. If the distribution of the p-values is uniform, this 
indicates that under the null hypothesis, our method will only reject the $H_0$
at a rate of $\alpha$.
A zoomed in plot for probabilities between 0 and 0.1 is
also provided, as
that is the most common range for significance levels $\alpha$.

\jy{Remove the white space in the pdf figures. Could put the zoomed version in
  the same figure.}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width = \textwidth]{figures/zoom_normal}
  \caption{A Q-Q plot of the p-values displayed in Figure~\ref{fig:normal} zoomed in to 
  probabilities between 0 and  0.1.}
  \label{fig:zoom_normal}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/zoom_gamma}
  \caption{A Q-Q plot of the p-values displayed in Figure~\ref{fig:gamma} zoomed in to 
  probabilities between 0 and 0.1.}
  \label{fig:zoom_gamma}
\end{figure}

From Figures~\ref{fig:normal} and \ref{fig:gamma}, we can observe that 
although the extent to which the p-values appear to be 
completely aligned with the line
in the Q-Q plots is dependent on the sample size and the level of serial 
dependence.
A small sample size like $n = 100$ or $200$ seems adequate when Kendall's
$\tau$ is lower than 0.25. For Kendall's $\tau \geq 0.5$, a sample larger than
$n = 200$ seems necessary for the p-values to be normally distributed. For
Kendall's $\tau \geq 0.75$, a sample larger than $n = 800$ appears to be 
necessary, as the p-values are not aligned with the line in the Q-Q plots, even
in Figures~\ref{fig:zoom_normal} and \ref{fig:zoom_gamma}. This
is not necessarily a cause for concern, as (for normal margins)
a Kendall's $\tau$ corresponds to
a $\phi$ of 0.924, which is very high. Additionally, performance generally seems
better for negative $\tau$ values versus positive $\tau$ values of the same
strength. Results are very similar for normal and gamma margins.
\jy{Summarize findings in bullet points before forming text.}

Under $H_0$, the p-values for the non-parametric block bootstrap KS test are
\jy{Not true from the plots.}
uniformly distributed. This is an indication that our method 
holds its size and that we can expect the probability
of a type 1 error to be the significance level $\alpha$.

\subsection{Power}
The power of the proposed test is investigated with data generated from
distributions that are not in the hypothesized distribution family. If the test
is indeed powerful, we would ideally want $\beta$ \jy{beta has been
  used. Discuss in terms of power: the probability of rejecting $H_0$ when $H_0$
  is false.}
to be 0, but we
expect it to be generally low. Again, we must try the method with different
marginal distributions to ensure that it is robust.
We would then expect the distribution of p-values to be non-uniform, and the 
rate \jy{Does not sound right. What is $\alpha$?}
of $\#P < \alpha$ to be high, meaning a higher density of low p-values.

\jy{Description too wordy, Be concise. Basically generate data from one
  distribution but test for the other distribution.}
Like we did to see if the test holds its size, we generated time series with 
marginal distributions $N(8, 8)$ and 
$\Gamma(8, 1)$, Kendall's $\tau \in \{-.75, -.5, -.25, 0, .25, .5, .75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. However, 
to evaluate the test's power, when $X_t \sim N(8, 8)$, we tested
that the marginal distribution family is from the Gamma family,
or
$X_t \sim \Gamma(\cdot \mid \alpha, \beta), \quad t = 1, \ldots, n$.
for some unspecified $\alpha$ and $\beta$. Because the support of $N(8, 8)$ is
$(-\infty, \infty)$, but the support of $\Gamma(8, 1)$ is $(0, \infty)$, we used
the \textsl{truncdist} package \citep{truncdist} to truncate the series at 
values less than or equal to 0.
When $X_t \sim \Gamma(8, 1)$, we tested that the 
marginal distribution is from the
Normal family, or 
$X_t \sim N(\cdot \mid \mu, \sigma^2), \quad t = 1, \ldots, n$
for some unspecified $\mu$ and $\sigma$. For the block bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method $1000$ times 
to obtain a p-value $p_r$ for each $r \in \{1, \ldots, 1000\}$.

\jy{alpha has been used as gamma parameters. This is not necessary, Just say we
  checked the power when the significance level is .05.}
Given some significance level $\alpha$, we can compute a rejection rate 
$q = \#\{p_r < \alpha\} / 1000$ for $r \in \{1, \ldots, 1000\}$.
Additionally, we can construct a 95\% confidence interval for 
this rate. Below are tables containing rejection rates showcasing the 
differences when simulation settings are changed.

@@

Figure~\ref{fig:rr} displays the empirical power curve of the test at
significance level 0.05 as the lag-1 Kendall's~$\tau$ increases from
$-0.9$ to $0.9$. Using a sample size $n \geq 800$ appears to be sufficient
\jy{What are the findings? summarize bullet points before forming texts.}
when $\tau$ is not too high, but for $\tau \geq 0.75$, all sample sizes 
observed appeared to have less than desirable rejection rates. Results
do not appear to be to different when comparing gamma and normal 
margins, indicating that the method is robust to the marginal distribution.
The rejection rates for $n = 800$ are very 
high (almost 1), indicating that under $H_A$, the test is very powerful 
(low probability of failed rejection $\beta$) when a large enough
sample size is used.

\section{Real Data Examples}
\label{sec:real}

Using the \textsl{tseries} package  \citep{tseries}, 
we gathered daily closing prices for the S\&P 500 index from January 1st, 2020
to December 31st, 2023. Daily returns were obtained by taking the difference in
the logarithm of the closing prices.
Using our method, we tested that the daily returns are normally
distributed and Student's $t$ distributed 
with degrees of freedom $v \in \{30, 20, 10, 5, 4, 3, 2, 1\}$.
We tested the same hypotheses using \citet{babu2004goodness}'s 
non-parametric basic bootstrap bias correction and 
\citet{zeimbekakis2022misuses}'s parametric bootstrap method.
We used $B = 10,000$ bootstrap 
replicates in this analysis.
The p-values for these tests are highlighted in 
Table~\ref{table:SP5004}.

\jy{Put them into one table; add results for paremetric bootstrap with
  Zeimbakakis' approach with serial dependence accounted for; label the methods
  as serial dependence accounted for and not accounted for, two methods
  (parametric and nonparametric) under each.}

\jy{Discuss the difference between the semiparametric bootstrap and the proposed
  nonparametric bootstrap. The semiparametric method depends on a working copula
  model for the serial dependence. If that is severely misspecified, I'd expect
  poor performance. The proposed method does not need to specify that dependence
  structure.}

@@

\jy{Stock returns are known to be serially dependent and heavy-tailed. Search
  google scholar ``stylized facts of daily returns series'' and cite them.}

While differences in p-values among different methods are expected, our analysis
reveals numerous instances where conclusions at the 0.05 significance level
using our method differ from those obtained using the methods proposed by
\citet{babu2004goodness} and \citet{zeimbekakis2022misuses}.
\jy{I don't like one-sentence paragraphs.}

@@

The same analyses were repeated on the daily returns of the S\&P 100 index.
Our method demonstrates more instances where it reaches
different conclusions at the 0.05 significance level compared to other methods.
Our method does not reject the hypothesis that the series conforms to a
Student's $t$ distribution with degrees of freedom $v = 5$, yielding a p-value
of 0.0873. In contrast, \citet{babu2004goodness}'s method yields a p-value of 
0.0368, and \citet{zeimbekakis2022misuses}'s method yields a p-value of 0.03.
conclusions regarding the marginal distribution of the population under scrutin.

@@

We have also 
demonstrated possible applications of this method to financial data.
\jy{don't astrophie citet; let alone citep}
Through comparison with \citep{babu2004goodness}'s non-parametric
bootstrap bias correction and \citet{zeimbekakis2022misuses}'s parametric
bootstrap method, we have underscored the importance of accounting for serial
dependence in choosing an appropriate methodology.
This method holds relevance for studies aiming to assess if a serially dependent
time series adheres to a hypothesized distribution in scenarios where parameters
are unknown or unspecified. Future studies could investigate
\jy{Too vague.}
the method's performance on different marginal distributions and
different dependence structures. Additionally, one could propose a 
similar method
for comparing two serially dependent methods. One could also apply the method
showcased
in this paper to fields like earthquake prediction and astronomy.

\jy{summarize remarks in bullet points first. Let's talk on Tuesday.}
